---
title: "Eva's #30DayChartChallenge Gallery"
---

::: {.callout-note}
## #30DayChartChallenge

#30DayChartChallenge is a global, community-driven data visualization initiative taking place annually in April. Everyone is invited to create and share one chart per day for 30 days, each based on a different prompt.
:::

```{r}
#| label: session-setup

library(dplyr)
library(tidyr)
library(stringr)
```


```{r}
#| label: custom-functions

#' CleanPromptsMd
#' @param prompts_path - a character vector - path to a md file with categories and prompts per day as obtained from the official challenge repo
#' @returns A data frame with category and prompt and path to the produced plot for each day which was worked on

CleanPromptsMd <- function(prompts_path) {
  prompts <- read.delim(prompts_path, header = T, col.names = "prompt")
  
  # Remove markdown styling characters
  prompts <- prompts |>
    mutate(prompt = gsub("  |[\\]|[*]", "", prompt))

  # Isolate category amd prompt into separate columns
  prompts <- prompts |>
    mutate(
      prompt_day = as.numeric(gsub(". ", "", str_extract(string = prompt, pattern = "^\\d+. "))),
      category = case_when(
        is.na(prompt_day) ~ gsub(" ", "", prompt),
        TRUE ~ NA
      ),
      prompt = str_to_sentence(gsub("\\d+. ", "", prompt))
    ) |>
    fill(category, .direction = "down") |>
    filter_out(is.na(prompt_day))
  
  return(prompts)
}

#' DisplayPlots
#' @param df_plots - a data frame with columns category, prompt and path (figure) and path_script as obtained with the CleanPromptsMd() function
#' @returns A quarto syntax to display plots sorted by category and in a two-column layout.
#' Use in a code chunk with option `results: asis`

DisplayPlots <- function(df_plots) {
  for (ctg in categories) {
    ctg_plots <- filter(df_plots, category == ctg)

    cat(paste0(sprintf("## %s", ctg)), "\n\n")

    for (d in ctg_plots$prompt_day) {
      pprompt <- ctg_plots[ctg_plots$prompt_day == d, ]$prompt
      path <- ctg_plots[ctg_plots$prompt_day == d, ]$path
      path_script <- ctg_plots[ctg_plots$prompt_day == d, ]$path_script

      cat("::: {.panel-tabset}\n\n")

      cat(sprintf("### Day %s: %s\n\n", d, pprompt))

      cat(
        sprintf(
          '![Day %s: %s](%s){width="650px"}\n\n',
          d,
          pprompt,
          path
        )
      )

      cat("### Code\n\n")

      # For proper monospace formatting
      cat("```r {eval=FALSE}\n")
      cat(readLines(path_script), sep = "\n")
      cat("\n```\n\n")


      cat(":::\n\n") # panel-tabset end
    }
  }
}
```


# 2026

```{r}
#| label: extract-prompts-2026

prompts <- CleanPromptsMd(prompts_path = "../2026/2026-list_prompts.md")
```


```{r}
#| label: identify-prompts-with-plots-2026

plot_files <- list.files("../2026/plots/", pattern = "png", full.names = T)

df_plots <- data.frame(
  "path" = plot_files,
  "prompt_day" = as.numeric(str_extract(string = basename(plot_files), pattern = "^\\d+"))
) |> 
  left_join(prompts, by = "prompt_day")

df_plots <- df_plots |> 
  filter_out(grepl("exclude.png",  path))

# Categories with plots
categories <- unique(df_plots$category)

script_files <- list.files("../2026/", pattern = ".R$", full.names = T)
df_plots <- data.frame(
  "path_script" = script_files,
  "prompt_day" = as.numeric(str_extract(string = basename(script_files), pattern = "^\\d+"))
) |> 
  left_join(df_plots, by = "prompt_day")
```


```{r}
#| label: display-2026
#| results: asis

DisplayPlots(df_plots)
```


## Resources
- data from `{NHANES}` package for R  
- `PrecipGL` data from the `{Kendall}` package for R  
- population growth data by [UN(]https://data.un.org/) (file `SYB61_253_Population_Growth_Rates_in Urban areas and Capital cities.csv`)  

# 2025

```{r}
#| label: extract-prompts-2025

prompts <- CleanPromptsMd("../2025/2025-list_prompts.md")
```


```{r}
#| label: identify-prompts-with-plots-2025

plot_files <- list.files("../2025/plots/", pattern = "png", full.names = T)

df_plots <- data.frame(
  "path" = plot_files,
  "prompt_day" = as.numeric(str_extract(string = basename(plot_files), pattern = "^\\d+"))
) |> 
  left_join(prompts, by = "prompt_day")

df_plots <- df_plots |> 
  filter_out(grepl("exclude.png",  path))

# Categories with plots
categories <- unique(df_plots$category)

script_files <- list.files("../2025/", pattern = ".R$", full.names = T)
df_plots <- data.frame(
  "path_script" = script_files,
  "prompt_day" = as.numeric(str_extract(string = basename(script_files), pattern = "^\\d+"))
) |> 
  left_join(df_plots, by = "prompt_day")
```


```{r}
#| label: display-2025
#| results: asis

DisplayPlots(df_plots)
```


## Resources
- Our Wolrd in Data: [https://ourworldindata.org/grapher/number-of-internet-users?tab=table](https://ourworldindata.org/grapher/number-of-internet-users?tab=table)  
- Our Wolrd in Data: [https://ourworldindata.org/how-urban-is-the-world](https://ourworldindata.org/how-urban-is-the-world)  
- [Fisheries Inventory, Trout Data, at Rocky Mountain National Park 2021-2022 by data.gov](https://catalog.data.gov/dataset/fisheries-inventory-trout-data-at-rocky-mountain-national-park-2021-2022-open-format-datas)  
- [GHS-COUNTRY-STATS R2024A - GHSL Country Statistics by Degree of Urbanization, multitemporal (1975-2030)](https://data.jrc.ec.europa.eu/dataset/341c0608-5ca5-4ddb-b068-a412e35a3326)  
